---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Topic 1

### Exercise 

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Functions

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Viewing Dataset Dimensions

`dim()`, `nrow()`, `ncol()` are all useful for seeing the size of a dataset.

```{r quiz-dim}
quiz(
  question("How many rows are in the mtcars dataset?",
    answer("11"),
    answer("32", correct = TRUE),
    answer("472")
  ),
  question("How many columns are in the mtcars dataset?",
    answer("11", correct = TRUE),
    answer("32"),
    answer("472")
  ),
  question("What does `dim(mtcars)` return?",
    answer("number of of cells in dataset"),
    answer("number of columns, number of rows"),
    answer("number of rows, number of columns", correct = TRUE)
  )
  )
```

## Viewing Rows


Use a function to view the *first* 5 rows of `mtcars`

```{r print-limit-head, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-head-hint}
head(mtcars)
```

Use a function to view the *last* 5 rows of `mtcars`

```{r print-limit-tail, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-tail-hint}
tail(mtcars)
```

Modify the following code to view rows 6 - 9:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
mtcars[6:9,]
```


## Viewing Full Datasets


```{r quiz-full-dataset}
quiz(
  question("What should you type into the console to look at the mtcars dataset in its own tab?",
    answer("mtcars"),
    answer("View(mtcars)", correct = TRUE),
    answer("head(mtcars)")
  ),
  question("What should you type into the console to look at the full mtcars dataset in the console itself?",
    answer("mtcars", correct = TRUE),
    answer("View(mtcars)"),
    answer("head(mtcars)")
  ),
  question("What should you type into the console to quickly look at the range of values for each column in mtcars?",
    answer("mtcars[,1:3]"),
    answer("range(mtcars)"),
    answer("summary(mtcars)", correct = TRUE)
  )
)
```



## Viewing Columns

Return the names of all columns in `mtcars`

```{r print-names, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-names-hint}
names(mtcars)
```

Return the *first three* columns in `mtcars`

```{r print-limit-cols, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-cols-hint}
mtcars[,1:3]
```

Return just the *gear* column from `mtcars`

```{r print-gear, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-gear-hint}
mtcars$gear
```